\begin{center}
    \vspace*{0.5em}
    {\LARGE\heiti\bfseries 跨模态抗污染异常检测的统一基准评测与主动去噪防御研究 \par}
    \vspace{0.6em}
    {\normalsize 杨景翔\,(2351576) \quad 徐云鹏\,(2353583) \quad 陈艺龙\,(2352359) \quad 李雪菲\,(2354093) \par}
    \vspace{1.0em}
\end{center}

\begin{minipage}{\textwidth}
\small
\noindent{\heiti\bfseries 摘\quad 要：}\;
异常检测旨在从高维数据中识别偏离正常分布的稀有样本，是数据挖掘与智能诊断系统的关键环节。然而现有评测体系普遍存在两点局限：一是\textbf{评测口径碎片化}，表格、时序、图、图像与文本各模态的基准相互割裂，难以在统一标准下横向比较；二是\textbf{对标签污染缺乏系统刻画}，而真实标注中的标签噪声会显著破坏监督模型的决策边界。针对上述问题，本文构建了一个跨越五大模态的统一抗污染异常检测基准平台，在 29 个真实数据集上系统评测了 26 种经典与前沿算法，并设计了一套即插即用的主动去噪防御包装器。我们通过四组递进实验，依次刻画了算法在无噪状态下的性能上限（Exp1）、在 0\%--20\% 对称标签翻转下的退化轨迹（Exp2）、跨模态的相对分类难度（Exp3），以及主动防御在 20\% 重度污染下的去噪挽救能力与失效边界（Exp4）。实验表明：以四分位距统计为清洁器、以样本剪裁为策略的 \texttt{IQR\_Trim} 防御对高容量监督模型保护效果最佳，使多层感知机的 AUC-ROC 绝对回升 \textbf{2.52} 个百分点，且剪裁策略在各类清洁器上均稳定优于标签纠正策略；与此同时，该防御对线性逻辑回归与表格基础模型 TabPFN 产生轻微负迁移，从经验上印证了去噪鲁棒学习中的“没有免费的午餐”原则。本研究为跨模态异常检测的鲁棒性评估提供了统一底座，并为工业级抗污染系统的部署给出了可操作的策略依据。\par
\vspace{0.4em}
\noindent{\heiti\bfseries 关键词：}\;异常检测；\;跨模态基准；\;标签污染；\;鲁棒学习；\;主动去噪防御；\;失效边界\par

\vspace{0.6em}
\noindent{\heiti\bfseries Abstract:}\;
{\itshape Anomaly detection identifies rare, deviating samples in high-dimensional data and underpins many data-mining and intelligent-diagnosis systems. Existing evaluation practice, however, is fragmented across modalities and rarely models the label contamination that pervades real annotation. We present a unified, contamination-aware benchmark spanning tabular, time-series, graph, image and text modalities, evaluating 26 algorithms over 29 real-world datasets, together with a plug-and-play active denoising defense. Through four progressive experiments we characterize clean-data performance ceilings, degradation trajectories under 0\%--20\% symmetric label flips, cross-modal relative difficulty, and the rescue effect and failure boundary of the defense under heavy contamination. An IQR-based sample-trimming defense (\texttt{IQR\_Trim}) restores up to \textbf{2.52} AUC-ROC points for an MLP under 20\% noise and consistently beats label correction, while inducing mild negative transfer on logistic regression and TabPFN---an empirical instance of the No-Free-Lunch principle in robust denoising.}\par
\vspace{0.4em}
\noindent{\heiti\bfseries Keywords:}\;{\itshape Anomaly Detection; Cross-Modal Benchmark; Label Contamination; Robust Learning; Active Denoising Defense; Failure Boundary}\par
\end{minipage}

\vspace{1.2em}
